---
title: "Merge our grid with NEFSC"
output: html_document
---

# Instructions

Would it be possible for you to add a column that gives the NOAA-NMFS Statistical Area number for each row in the data file?  

# Data

## NEFSC Shapefile

Shapefile provided by Chris on the NEFSC website:  ftp://ftp.nefsc.noaa.gov/pub/gis/

- Statistical_Areas_2010_withNames.shp
- Statistical_Areas_2010.shp

- The description and metadata for the shape file are here: https://www.fisheries.noaa.gov/inport/item/26262

This is the NEFSC-recommended citation for the shapefile:
 
Northeast Fisheries Science Center, 2021: NEFSC Statistical Areas from 2010-06-15 to 2010-08-15. NOAA National Centers for Environmental Information, https://www.fisheries.noaa.gov/inport/item/26262.

## Load Data

```{r load_data}

# oad our grid SF
grid_sf <- st_read(my_path("D","Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>% 
  select(index,geometry)


#Load NEFSC shapefile
nefsc_sf <- st_read(my_path("D","Spatial/nefsc", name = "Statistical_Areas_2010.shp")) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

# Visualize map (ok)
# nefsc_sf %>% 
#   st_simplify(preserveTopology = TRUE, dTolerance = 0.1) %>% 
#   ggplot() +
  # geom_sf()

```

## Merge shapefiles

```{r merge_shapefiles}

# Merge both sf
nefsc_index <- nefsc_sf %>% 
   st_join(grid_sf,
           join = st_intersects) %>% 
   filter(!is.na(index)) %>% 
  as.data.frame() %>% 
  select(Id,index) %>% 
  distinct(.keep_all = T) %>% 
  mutate(plop = "yes") %>% 
  left_join(my_path("D", "Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T)) %>% 
  select(Id,index,state)
 
# Make sure it works
nefsc_sf %>%
  left_join(nefsc_index) %>%
  filter(!is.na(index)) %>%
  ggplot() +
  geom_sf(aes(fill = index))


```

## Save shapefile

```{r}

# File name
file_name <- my_path("D","Spatial", name = "nefsc_index.csv")  

write_csv(nefsc_index %>% select(-state), file_name)
```

